### 2014-07-28 paper/referee/response-revision1.txt
minor rev1 update

MN-14-1011-MJ: Authors' response to the reviewer's comments
===========================================================

We are grateful to the referee for the insightful feedback, which has helped us
to improve the presentation of the data products.  In his/her report, the 
referee identified areas where additional information would (i) strengthen the 
role of the paper as a user's guide to the survey, and (ii) would lessen the 
burden on the reader to track down information from previous works.  
We appreciate the merits of the referee's suggestions in this respect, and have 
acted upon them by adding a total of N new pages, 1 new table, and 1 new figure 
to the manuscript.  These changes have been marked in bold in the manuscript.  
We have also added N new sections of auxillary material to our web site as a 
direct consequence.

In what follows we will start by explaining our response to each of the issues 
which the referee identified as high-level remarks (9 items), followed by the
referee's additional comments which affect specific sections (8 items).

High-level items
----------------

[Referee]
> 1. This manuscript presents a large amount of information about IPHAS and the preparation
> of and contents of the data products.  It also relies heavily on previous
> publications for general descriptions of the survey, particulars of the instrumentation,
> data processing algorithms and earlier data releases.   As such, it puts considerable burden 
> on the readers to track down information from many different sources.  It is
> understandable that the authors prefer to not duplicate information that is already
> available elsewhere, but it will make it much easier to understand and use the IPHAS data
> products if at least some of the general information  could be repeated in this
> manuscript.
>
> For example, it would be helpful to include concise summary of the IPHAS survey
> parameters and this data release in tabular form.  This summary would serve as
> a "IPHAS-At-A-Glance" for users, and coule include the survey area, dates, telescope,
> camera and survey bandpasses, etc.  Much of the information is presented in the paper,
> but widely spread over the manuscript.  Without at least a guide to finding the
> information, users will be challenged to find what they need.

[Authors' response]
We have kept the referee's recommendation to lessen the burden on the reader in
mind while revising the paper in response to the items to follow below.
As request by the referee, we have included a concise summary table of the
survey parameters at the start of Section 2, in which the 15 key properties
of the survey are summarised (Table 1).

[Referee]
> 2. It was not immediately obvious if the images and catalogue from the Second IPHAS
> release supersede those in the 2008 Initial Data Release.  I believe the
> Second Release includes a reprocessing of all of the survey data to-date, so it
> is intended to replace the Initial Release, but that should be stated explicitly
> in the Introduction (e.g. in the 3rd paragraph).
>
> If my assumption is correct, there should be a summary statement about why the
> Second Release images and catalogue are in fact better products.  For example,
> better calibrations, processing algorithms, etc.

[Authors' response]
The referee correctly identified that we forgot to state that this new release 
supersedes the previous one.  We have added an explicit statement in this sense
to the third paragraph of Section 1, and in addition we have edited the last
paragraph of Section 3.1 to list the key differences between the past and the
present release.

[Referee]
> 3. As stated in Section 7.2, the aim of IPHAS is to enable the discovery of
> emission line objects in the Galactic Plane.  Consequently, I was very surprised
> that no discussion of the sensitivity to line emission (e.g. H-alpha EW) was
> presented in the manuscript.  Perhaps this will be addressed further in the planned
> Barentsen et al paper, but it seems that at least some rudimentary demonstration
> of the sensitivity to line emission is a requirement for this paper.  The
> general photometric sensitivity in the three bands is discussed in section 6.2, and
> an illustration is given how to find emission-line candidates using the color-color
> diagram in Figure 18.  I strongly recommend that at least some examples be given
> of the r-H-alpha color excess as a function of line EW for a sample of objects
> with published spectra.

[Authors' response]
The referee requested a discussion on the sensitivity of the H-alpha 
photometry to line emission to be included in this work.  In response, we have
added one new section (Sect 7.3) and one new figure (Fig 19) to the paper.
In this new section, which is titled "r-Halpha excess as a quantitative measure
of Halpha emission", we compare the spectroscopic H-alpha EWs for a sample of
classical Be stars from Raddi et al (2014) against the position of these
objects in the IPHAS colour-colour diagram.  A good agreement is found.
Moreover, the accompanying discussion offers a brief guide to the previous 
works on this topic.

[Referee]
> 4. Many of the successful large surveys include a list of known caveats concerning
> released data products to help user's avoid pitfalls in their research.  I suggest that
> the authors generate a short list of the most important things about which users
> should be aware when using IPHAS data.  Users appreciate knowing how to avoid mistakes
> due to misinterpretation of known issues, and do not think less of a product because
> of them.

[Authors' response]
In response to the referee's suggestion, we have (i) introduced a summary of
the most common pitfalls at the start of Section 6, and (ii) have added a
caveats section to our data release web page (www.iphas.org/dr2).
We note that it is not possible, at this stage, to include an exhaustive list
of all the caveats, because we cannot anticipate all forms of use of the data.
Hence in Section 6 we alert the reader to the fact that we offer a caveats
section on the website, which will be updated as user experience accumulates.

[Referee]
> 5. In general, the images do not receive the same level of description in the manuscript
> as does the Catalogue.  Any guidance that can be given to users about them would be
> beneficial.  For example, at the very least include a description of the FITS image
> headers, analogous to the descriptions of the catalog columns that is given in
> the Appendix.  They should also mention that tables containing the image metadata are available at IPHAS web site.

[Authors' response]
Although the paper intended to focus on the catalogue product, we agree with the
referee that it would be beneficial to give users additional guidance on the use 
of the image products.  For this reason, we have made significant improvements 
to our web site by (i) providing a query tool to retrieve IPHAS images by 
coordinates, and (ii) provide more comprehensive documentation on the metadata, 
the FITS header keywords, and the flux calibration (see www.iphas.org/images).
Section 6 of the paper has been edited to advertise these resources, which are 
set to be updated in the future whilst we prepare a data release which focuses
more strongly on image and image mosaic products.

[Referee]
> 6. The fact that only 39% of the catalogue entries satisfy the criteria necessary to be
> classified as "reliable" begs the question "just what are the other 61% of objects."
> The authors should add more information about the general reliability of the
> catalogue with and without their criteria. For example, If a users selects
> "reliable" category sources, what should they expect the net reliability to be?
> Certainly not 100%.  How does reliability vary with source magnitude, SNR, or other 
> measurement parameters?  How much does the reliability vary with source environment
> (e.g. crowding, complex backgrounds)?
>
> Reliability can be gauged in a number of ways, such as comparisons with external,
> deeper catalogues (e.g. SDSS, UKIDSS GPS).

[Authors' response]
The referee questioned the nature of the objects which are marked as "reliable"
in the catalogue, and in particular questioned the nature of the large number of 
sources not flagged as "reliable".  Having considered this comment, we have come 
to conclude that we made a poor choice of terminology by adopting the labels 
"reliable" and "veryReliable" as quality qualifiers.
These classes, which are based on stringent quality criteria set out in 
Tables 6 and 7, aim to help users identify high-quality objects, e.g. by 
demanding a 10-sigma detection in all bands.  It goes without saying that 
there are plenty of objects which do not strictly satisfy these criteria but are 
nevertheless genuine astrophysical objects.  The qualifier "reliable" might 
however confer the incorrect impression that such sources are "unreliable".
As a result, we have decided to change the terminology by renaming the class 
previously known as "reliable" into "a10" (which is short for "all-band 10-sigma"), 
and renaming "veryReliable" into "a10point".  We believe this change in 
terminology will better serve as an operational description of the detection quality.

In addition to this change in terminology, we have followed the referee's suggestion
to gauge the reliability of our objects against an external catalogue.  That is,
we have introduced a new Section 6.2, titled "Source reliability", in which the 
reliability of IPHAS sources as a function of the quality criteria is gauged 
by comparing out catalogue against SDSS.  Moreover, we note that a separate paper 
is in an advanced state of preparation (Farnhill et al.), which will investigate 
the issues surrounding the completeness and reliability of our catalogue, 
e.g. by injecting artificial sources into the images and measuring their recovery 
rates, whilst also comparing source counts between SDSS and IPHAS in more detail.

[Referee]
> 7. The authors have done a good job describing the photometric measurements and
> calibration accuracy.  However, there is relatively little discussion of the
> astrometric position reconstruction other than saying it is done relative to
> the 2MASS Point Source Catalogue, and there is no validation of the astrometric
> quality of the Catalogue other than to mention matching success.  Some quantification
> of the astrometric performance should be added (e.g. mean position residuals as a function
> of magnitude from comparisons with the astrometric reference stars, or external
> astrometric collections, such as UCAC4).
>
> Positions in the IPHAS Catalogue have been reconstructed with respect to 2MASS.
> However, the authors do not state if the proper motion of the 2MASS reference stars 
> was taken into account.  Although small, the net motion of the reference stars can
> collectively imprint on the Catalogue because the IPHAS observations were made at
> different epochs than 2MASS, in some cases as much as 12 years later.   
> Failure to do so can imprint the proper motion onto IPHAS, and also because the 
> IPHAS measurements span nearly 10 years (2003-2012), there can be secular variations.

[Authors' response]
astrometry ...
proper motions ...

(There does not appear to be an item number 8 in the reviewer's comments.)

[Referee]
> 9. The authors point out in the introduction that IPHAS will complement a number of
> new infrared surveys of the Galactic Plane.  The paper would benefit from at least
> one cross-comparison with the catalogue of one of those surveys, such as GLIMPSE
> or MIPSGAL, as an illustration of how the combination of surveys will enable
> new science.  At present, the only tie-in to the infrared surveys is the
> visual comparison between IPHAS and GLIMPSE images of one field.

[Authors' response]
In response to the referee's suggestion to include a cross-comparison between
IPHAS and an infrared survey, we have added a brief discussion on the overlap
between the GLIMPSE and IPHAS catalogues at the end of Section 7.2.  The 
discussion warns the reader that an optical survey will naturally be 
extinction-limited.

Section-specific remarks
------------------------

[Referee]
> 1. In Section 3.1, paragraph 1, sentence 1:
> "archival" should be "archiving"

[Authors' response]
This has been corrected.

[Referee]
> 2. In the second paragraph of section 3.1, it is stated that a deblending routine
> is used to remove contaminating sources.  Additional information about this routine
> should be provided.  Are contaminants removed from the sky background measurement,
> the source measurement aperture, both?

[Referee]
> 3. At the beginning of the Data Reduction section (section 3), it would be helpful
> to provide a summary of how data processing for this release differed from that
> used for the Initial Data Release.

[Referee]
> 4.a. It is stated in section 3.2 that the exposures were discarded unless they
> had 5-sigma limiting magnitudes of r>20 or i>19 and H-apha>19 mag. How were the
> limiting magnitudes for an exposure estimated?

[Referee]
> 4.b. Section 3.2 also specifies that an ellipticity limit of eps>0.3 was used.
> What is the justification of this limit?  For example, what is the potential impact on
> photometry and sensitivity at eps=0.25?

[Referee]
> 5.a. Were the atmospheric extinction coefficients of 0.09 (r) and 0.05 (i) simply
> adopted, or was there any attempt to measure them?  Is there an estimate of
> how much the extinction may have varied with time, either during a night, or
> seasonally?

[Referee]
> 5.b. The PERCORR factor in equation 1 is used to correct for "sporadic gain variation."
> What was the nature of the variations when non-zero?  How was it measured?
> What is the size of the correction when it was applied?

[Referee]
> 5.c. The first paragraph of section 4.2 on Global Calibration states that calibration
> errors  can be up to 0.1 magnitude in reference to the colour-colour diagram shown in
> Figure 4a. There are outlier populations in that figure that are up to 0.6 mags
> from the main locus of points.  These seem to be accurately corrected by
> the recalibration, but the text seems to understate the severity of the
> systematics.

[Referee]
> 5.d.  It is stated in section 4.2, paragraph 2 that 2MASS observed six calibration
> fields each hour.  That is not correct.  2MASS observed two calibration fields each hour.  However, each observation of a field consisted of six scans.

[Referee]
> 5.e The authors may wish to specify in Section 4.3 that the SDSS DR9 Galactic Plane data
> that were used to test the IPHAS calibration were specifically from the SEGUE program
> of SDSS II, and cite the appropriate reference for SEGUE.

[Referee]
> 6.a. Section 5.1
> Would it be possible to provide the aperture corrections applied to the aperture
> photometry for each image along with the image metadata?

> Shouldn't the "saturated" column flag be specified per band?  Surely it must
> be possible for extremely red objects to be saturated in the i-band exposure
> but not r-band.

> 6.b. Section 5.2
> Little information about the specifics of band-merging is provided other than the
> reference to Taylor 2006 and TMATCHN.  Users will be forced to track down
> the reference to understand the process.  The authors should consider giving a short
> summary of the algorithm.  Some important information that should be provided are:

- How is confusion handled in the band merging process (e.g. what happens when
more than one i-band detection can be matched to an r-band detection?).
Source confusion in the band-merging process is mentioned in the final
paragraph of section 6.3.

- Is there any band-filling when a source is not detected in all bands?  If so, what
is the algorithm used to provide the flux limits in the non-detected bands?

6.c. Section 5.3 - Table 4

It is not clear what is the specific value is of the average magnitude given for each of
the different band-detection combinations in Table 4?  Wouldn't something like the peak
of the log N - log M curves be of more interest to provide some sense of the
completeness and/or depth of the Catalogue?

6.d. Section 5.4

While the authors provide a thorough description of the criteria used for
Catalogue selection, no reasons were offered for why those criteria were adopted.
For example, were there requirements for completeness and reliability that were
being satisfies?  If not, what were the objectives?

In the first paragraph of section 5.4, it is stated the three criteria were
enforced to limit the size of the catalogue.  Was this the main driver for establishing
the top-level criteria?  It seems unusual in this day and age of massive catalogues
that this would be a driver.

As stated earlier in this review, some description should be offered as to the nature of
the very large number of detections in the catalogue that are not classified as "reliable."
It would be useful to describe them to the user, and to offer suggestions of
how to recognize problematic detections, and also ones that may be real.

7. Section 6.1, par. 2, sentence 1

Should the work "accuracy" be "reliability"?

8. Figure 11 - caption

The reference to the left and right panels should presumably be to the top and bottom panels,
unless the two figures will be arranged differently in the published version.